Promise and peril: Agnostic identification methods\\ for detecting differential item functioning

### 2020-03-19 paper/paper_1_intro.Rmd
getting there

Inspired by @camilli1992conceptual, we think of differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies DIF.

Contrarily, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study. They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is $$P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)$$ where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{ref}} = -1$ for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model, the probability student $i$ responds correctly to item $j$ is
$$
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
$$
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

>>>>

Following @camilli1992conceptual, we conceptualize differential item functioning (DIF) as a varying relationship between ability and probability of correct response for students from different groups that manifests when an item response model with too few ability dimensions is imposed. From this perspective, the term "differential item functioning" is, perhaps, a misnomer as DIF is better thought of as a property of the student, as opposed to the item. For example, @ackerman1992didactic describes a scenario in which a test intends to measure a student's math ability, but performance also depends on their verbal ability. In this case, math ability is the "target ability," and verbal ability is the "nuisance ability." Fitting a unidimensional item response model to this test results in students with low verbal ability receiving a score systematically lower than their true math ability; therein lies "DIF".

Despite this, the usual setup of DIF simulation studies frames DIF as a property of the item. For example, @kopf2015framework simulate students as belonging to either a reference or focal group. They fix the item easinesses for the reference group, ${b_j}^{\text{ref}}$, to values from a previous study. They set item easinesses for the focal group to ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for items without DIF and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.6$ for items with DIF, where 0.6 is the magnitude of DIF in logits. They simulate student ability ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. They generate item responses according to the Rasch model, which specifies that the probability of student $i$ responding correctly to item $j$ is
\begin{align}
P(y_{ij} = 1 | \theta_i, b_j) = \sigma(\theta_i + b_j)
\end{align}
where $\sigma(x) = \frac{e^x}{e^x + 1}$ is the standard logistic function.

<!-- future improvement would be 1) show that for some student they equal just to make it ultra clear and 2) note that there are infinite possibilities the key is that the second term is -0.6 for students from the focal group -->

For every DIF simulation study framed in terms of item parameters that vary across groups, there is a mathematically equivalent setup in which students' abilities are multidimensional. For example, to translate the @kopf2015framework simulation from the DIF-as-item-property view to the DIF-as-student-property view, item easiness is set to what was previously ${b_j}^{\text{ref}}$ for all students. Student ability is expanded to two dimensions, the target ability dimension and nuisance ability dimension. The target ability is what was previously just ability where ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. The nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ for all students in the reference group and ${\eta_i}^{\text{foc}} = -1$ for all students in the focal group. We now need to use a compensatory 2PL model where the slope on target ability $a_{1j} = 1$ for all items (consistent with the Rasch model) and the slope on nuisance ability $a_{2j} = 0.6$ for all items with DIF, and $a_{2j} = 0$ otherwise. Again, 0.6 is the magnitude of DIF in logits. According to the two-dimensional compensatory 2PL model [@thissen1986taxonomy], the probability student $i$ responds correctly to item $j$ is
\begin{align}
\text{Pr}(y_{ij} = 1 | \theta_i, \eta_i, a_{1j}, a_{2j}, b_j) = \sigma(a_{1j}\theta_i + a_{2j}\eta_i + b_j).
\end{align}
This translation between views makes explicit that nearly all DIF simulation studies have, perhaps suboptimally, examined the unrealistic scenario in which there is no variation in the nuisance ability for students in the same group.

### 2020-03-19 paper/paper_2a_anchor_items.Rmd
getting there

The unsolved and interesting problem is how to link groups in the common circumstance in which no a priori assumptions — about either relative group ability or which items might have DIF — can be made. That is, we're interested in agnostic identification methods (hereafter referred to as "AI methods") so that LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same [e.g. @kopf2015framework], and that we risk contributing to a proliferation of names.

In this section, we summarize existing AI methods and propose extensions of those methods. We use a simple, one-run simulation to demonstrate the methods: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{foc}} = -1$. The slope on target ability is set to $a_{1j} = 1$ for all items. The slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items.

In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively. We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$. The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$. The most common AI approach is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.

@@

@@

@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can they be sure that its those 10 items and not the other 11 items with DIF? They cannot be. Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group. Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor]. In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting]. A simple thought experiment illustrates how AOAA fails: Imagine a test with a sufficiently large number of students and three items where the first item has DIF, and the other two do not. Using AOAA, all items test positive for DIF. The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated. This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."

@@

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively. @kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse, not incorrect) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items test positive for DIF, too.

@@

@@

Table \ref{table:themethods} summarizes the three all-others-as-anchors methods. It's useful to remember that AOAA is not an iterative procedure. The methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method. In our one-run simulation, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

@@

\begin{table}[H]
\caption{Summary of the three all-others-as-anchors agnostic identification methods}
\centering
\begin{tabular}{|p{2.5cm}|p{6cm}|p{4cm}|}
\toprule

Method & Description & Literature \\

\midrule

AOAA & Test if each item has DIF by using all of the other items as anchors (not iterative). & Originally proposed by \cite{lord1980} and formalized by \cite{thissen1993detection} \\\hline

AOAA-AS & The first iteration is AOAA. All items that test positive for DIF are removed from the anchor set. Continue iterating until no new items test positive for DIF. & Proposed by \cite{drasgow1987study} \\\hline

AOAA-OAT & The first iteration is AOAA. Only the item that shows the most extreme DIF is removed from the anchor set. Continue iterating until no new items test positive for DIF. & To our knowledge, not proposed or used previously \\

\bottomrule
\end{tabular}
\label{table:themethods}
\end{table}

@@

@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters. For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$. The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p. 323].

@@

We propose a new cluster-based approach, which we call "equal means clustering" (EMC). Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$ of differences in item easiness, $\tilde{d_j} = \tilde{b_j}^\text{ref} - \tilde{b_j}^\text{foc}$, where the model is identified by setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, thus the name "equal means clustering"). As a result, all differences in performance — either from group ability differences or DIF — manifesting in the item easiness difference parameter $\tilde{d_j}$. The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.

Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating. The largest cluster is chosen as the anchor cluster. If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected. This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF. In our one-run simulation, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.

### The equal means, multiple imputation logit graph (EM-MILG)

AI methods are generally designed to automatically detect DIF without any human judgement. On the other hand, it can be useful to present information to the analyst in a way that empowers them to see DIF and potentially even select the anchor items.

We propose a new method, the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst. Like EMC, EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc}$ to $0$. Again, the result is that all differences in performance manifest in $\tilde{\mathbf{d}}$. To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical]. Then, multiple imputations (MI) [@yang2012characterizing] are drawn to estimate the distribution of $\tilde{d_j}$ for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our one-run simulation. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The challenge, then, is to select the anchor items. The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased and can be used as anchor items.

After choosing the anchor items, the model is refit. The new model is identified by setting $d_j = 0$ for the anchor items, instead of by setting $\mu^\text{foc} = 0$. The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the new model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our one-run simulation, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

One of our key concerns with typical AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly identified the correct anchor items. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF anytime a model is fit using anchor items, not just when the first step involves the EM-MILG. In particular, we could visualize DIF in the non-anchor items after using, for example, AOAA-OAT or EMC to identify the model.

>>>>

The unsolved problem that we address here is how to link groups in the typical setting in which no a prior assumptions — about either relative group ability or which items might have DIF — can be made. That is, we're interested in agnostic identification methods (hereafter referred to as "AI methods") so that LRTs can be used to test items for DIF. For simplicity, we focus on the Rasch model to isolate the fundamental issues in DIF detection. And, in search of a coherent framework, we sometimes edit names of existing methods. We recognize that others have done the same [e.g. @kopf2015framework], and that we risk contributing to a proliferation of names.

In this section, we summarize existing AI methods and propose a variety of extensions. We use a simple demonstration data to illustrate how these methods work: 10,000 reference group students and 10,000 focal group students taking an eight-item test. Target ability is simulated ${\theta_i}^{\text{ref}} \sim N(0,1)$ for students in the reference group and ${\theta_i}^{\text{foc}} \sim N(-1,1)$ for students in the focal group. Nuisance ability is set to ${\eta_i}^{\text{ref}} = 0$ and ${\eta_i}^{\text{foc}} = -1$. The slope on target ability is set to $a_{1j} = 1$ for all items. The slope on nuisance ability is set to $a_{2j} = 0.5$ for the last three items (the items with DIF) and $a_{2j} = 0$ otherwise. We can also, of course, describe these conditions from the DIF-as-item-property view where there is no nuisance ability. Instead, ${b_j}^{\text{foc}} = {b_j}^{\text{ref}}$ for the first five items and ${b_j}^{\text{foc}} = {b_j}^{\text{ref}} - 0.5$ for the last three items. To be sure, this data is used only for demonstrative purposes. We compare the performance of the AI methods in a more realistic scenario in the [the simulation study](#simstudy).

In general, we denote the mean ability of the reference and focal group as $\mu^\text{ref}$ and $\mu^\text{foc}$, respectively. We follow the common and inconsequential practice of identifying the scale by setting $\mu^\text{ref} = 0$. The fundamental challenge is to use an AI method to identify the model, which allows for disentagling the estimation of $\hat\mu^\text{foc}$ from the estimation of $\hat d = \hat{b_j}^{\text{foc}} - \hat{b_j}^{\text{ref}}$, the difference in item easiness. The most common AI approach is to select a group of anchor items that are assumed to be DIF-free. These anchor items identify the model, thereby allowing for the estimation of $\hat\mu^\text{foc}$, and the remaining items can be tested for DIF using an LRT.

@@

## The equal means, multiple imputation logit graph (EM-MILG) {#emmilg}

AI methods typically work like a black box. The analyst puts their item response data in and the black box outputs an identification assumption to be used when fitting the final model. These types of AI methods have their role — and the main focus of this paper is on them — but wouldn't it be nice if there was a way for the analyst to visualize the possibility of DIF in their data?

We think so and, in response, propose the "equal means, multiple imputation logit graph" (EM-MILG), which presents information about potential DIF to the analyst. EM-MILG begins by fitting a unidimensional Rasch model to the data that is identified by setting $\mu^\text{foc} = 0$. Recall that $\mu^\text{ref} = 0$, so the assumption is that the groups have equal mean ability. The result is that all differences in performance — either from group ability differences or DIF — manifest in the item easiness difference parameter $\tilde{d_j}$. The tilde is used to denote parameters estimated with $\mu^\text{foc}$ arbitrarily set to $0$.

To measure the variation in each $\tilde{{d_j}}$, the item parameter covariance matrix is estimated using Oakes' identity [@chalmers2018numerical]. Then, multiple imputations (MI) [@yang2012characterizing] are drawn to estimate the distribution of $\tilde{d_j}$ for each item. These are the distributions displayed in a EM-MILG. The method is inspired in part by @pohl2017cluster who fit a model with both the reference and focal group means set to 0 in a pedagogical example, and @talbot2013taking who fixed both pre-test and post-test means to 0 in order to estimate item-specific learning gains.

Figure \ref{fig:emmilg} shows the EM-MILG for our demonstration data. We cannot state strongly enough that the EM-MILG contains all possible information about the difference in group performance. The analyst might assume that — because there are five items where the reference group outperforms the focal group by approximately 1 logit and only three items where the difference is 1.5 logits — items 1-5 are unbiased. These unbiased items are known as anchor items. After anchor items have been selected, the model is refit where the assumption that $\mu^\text{foc} = 0$ is dropped and replaced with the assumption that the anchor items are DIF-free (i.e., $d_j = 0$).

The same process of using multiple imputations to estimate the distribution of $\tilde{d_j}$ can be used with the final model. Because the equal means assumption is not made, we refer to the resulting visualization as a multiple imputations logit graph (MILG). As expected, in our demonstration data, selecting the first five items as anchors correctly results in $\hat d_j \approx 0.5$ for the items with DIF as is shown in the MILG in Figure \ref{fig:milg}.

In general, one of our key concerns with traditional AI methods is that they can lull the analyst into a false sense of security. Too often, the analyst chooses a method, implements it, and then proceeds as if the method certainly resulted in a model that reflects some objective reality. A EM-MILG combats this concern by presenting all information clearly to the analyst. In the previous example, the analyst may be wary of their results, having seen how arbitrary it was to conclude that the first five and not the last three items are unbiased. Even when other AI methods are used, the analyst can use the EM-MILG as the first step in order to give them a sense of their item response data. And, of course, the MILG can be used to visualize DIF regardless of how the model is identified. We now move on to more traditional AI methods.

@@

@edelen2006identification used AOAA to look for DIF between the English and Spanish versions of the 21-item Mini-Mental State Examination and found that 10 of the 21 items exhibited DIF. How can they be sure that its those 10 items and not the other 11 items with DIF? We argue that this method cannot tell us with certainty that it is those 10 items that have DIF while the other 11 do not. Implicit in the use of AOAA is the assumption that all items not being tested do not exhibit DIF, which is, of course, impossible. More practically, it is thought that AOAA will perform well if a small minority of items have DIF or the DIF is balanced such that some items are biased against the focal group, while others are biased against the reference group. Undesirably, most applications of AI methods and many simulation studies do not make explicit the assumptions of the AI method [@strobl2018anchor]. In this way, psychometricians might benefit from adopting economists' habit of explicitly stating assumptions and debating their plausibility.

Researchers have noticed the circular logic of AOAA, but have mostly described it indirectly by pointing out inflated Type I errors in simulation studies [@stark2006detecting]. A simple thought experiment illustrates how AOAA fails: Imagine a three item test with a sufficiently large number of students where the first item has DIF, and the other two do not. Using AOAA, all items test positive for DIF. The last two items incorrectly test positive because including the first item in the anchor set causes the group ability difference to be misestimated. This phenomenon of items with real DIF inducing the appearance of DIF in other items was only indirectly discussed in the literature until @andrich2012real coined the term "artificial DIF."

@@

One way to attempt to counter artificial DIF is with purification through iterative anchor selection. For example, @drasgow1987study started with AOAA, removed items displaying DIF from the anchor set, then repeated the process iteratively — with items that have been removed from the anchor set allowed to have free parameters across groups in both the baseline and flexible model — until no more items tested positively. @kopf2015framework named this technique Iterative-backward-AOAA with "backward" (as in reverse) referring to beginning with the assumption that all items are DIF-free. We find it clearer to refer to this method as all-others-as-anchors-all-significiant (AOAA-AS). Appending all-significant indicates that anchor selection is made iteratively with all items that test positive for DIF being removed from the anchor set. AOAA-AS is implemented in the mirt R package, and is called by passing scheme = "drop_sequential" to the DIF function.

AOAA-AS might seem like an improvement, but it doesn't solve a fundamental problem of AOAA: What does one do when all items test positive for DIF? With a sufficient sample size and at least one item with DIF, this will necessarily be the case. In our thought experiment, we get the same result with AOAA-AS as we did with AOAA: All items test positive for DIF, and there are no anchor items. @kopf2015framework encountered precisely this problem in their simulation study and chose to select a single anchor item randomly. @woods2009empirical suggested a more straightforward, one-step method which uses AOAA and selects the, say, four items that exhibit the least amount of DIF. It's unclear if one should proceed if those four items also test positive for DIF.

@@

<!-- improvement would be to explain what more homogeneous means in more detail -->

@@

@@

It's useful to remember that AOAA is not an iterative procedure. In constrast, the methods with a hyphen, AOAA-OAT and AAOA-AS, are iterative procedures with AOAA-OAT being the new, more conservative method. All of the methods described in this paper are summarized in Table \ref{table:allmethods}. For our demonstration data, AOAA and AOAA-AS both failed to select any anchor items, leaving the model unidentified. On the other hand, AOAA-OAT worked perfectly, selecting exactly the first five items as anchor items.

<!-- probably should add transition here -->

@@

<!-- ben d rightly points out that it would be good to say more what we mean by "similarly" - could look at how they describe it -->

@bechger2015statistical proposed selecting anchor items by identifying clusters of items that function similarly and then choosing one of those clusters to be the "anchor cluster." They pointed out that one way around the unidentifiability issue is to consider only relative item parameters. For each group, the relative easinesses for each pair of items can be stored in the matrix $\mathbf{R}^{\text{ref}}$ with entries ${R_{xy}}^{\text{ref}} = {b_x}^{\text{ref}} - {b_y}^{\text{ref}}$, where $x$ and $y$ both index the items. The ultimate matrix of interest is $\Delta \mathbf{R} \equiv \mathbf{R}^{\text{ref}}-\mathbf{R}^{\text{foc}}$ which is the "differences between groups in the pairwise differences in (easiness) between items" [p. 323].

@@

We propose a new cluster-based approach, which we call "equal means clustering" (EMC). Instead of working with an arbitrary column from $\Delta \mathbf{R}$, we work with the vector $\tilde{\mathbf{d}}$, which, as described in the [EM-MILG section](#emmilg), comes from setting $\mu^\text{foc} = 0$ (recall $\mu^\text{ref}$ is always set to $0$, and thus the name "equal means clustering").

Instead of choosing k with BIC, we use the gap statistic method recommended by @hastie2001estimating. The largest cluster is chosen as the anchor cluster. If there is a tie for the largest cluster, the cluster with the lowest standard deviation of $\tilde{\mathbf{d}}$ is selected. This process assumes that the largest cluster (but not necessarily the majority) of items do not contain DIF. For our demonstration data, this approach worked perfectly by finding two clusters of items, one corresponding to the anchor items and the other corresponding to the items with DIF.
